Index measures for oak decline severity using phenotypic descriptors

### 2019-12-10 manuscript/manuscript.Rmd
added to introduction

- define phenotype and descriptor
- Acute and chronic decline associated phenotypes
- The oak decline spectrum move away from simplistic binary categorisations
- Manual assignment of decline status difficult prone to error for due to contribution of many factors. 
- Random forest
- Unsupervised learning for complex patterns between descriptors.

Oak decline syndromes are facilitated by a variety of both abiotic and biotic factors.

Abiotic factors can include...
Biotic factors can include...

As a results of this, there are many visual cues measures that contribute to what can be defined as a declining oak tree.
A severely declining oak is prominantly characterised by very poor crown condition and numerous dead branches.
Compounding this complexity is the addition of different types of oak decline (AOD) that are often categorised based on the duration over which a tree has been declining or the biotic agents/factors associated with a declining tree.
For instance a chronically declining oak can subsequently begin to also develop the stem bleeds associated with AOD.
Accurately diagnosing in the field the severity and type of oak decline present is difficult and can requires much experience, expertise and training.

Machine learning is finding increasing utility in biological and geographical applications due to it's...
A popular algroith is that of random forest due to it's versatility, relatively trivial training requirements and resiliance against overfitting [@breiman2001] (refs). 
It is an ensemble machine learning method where forests of bootstrap sampled decision trees are grown, yielding both out of bag (OOB) sample prediction accuracies and variable importance.
Unsupervised learning is also possible

@@

### Phenotypic descriptor preparation and analysis

@@

>>>>

- Increasing importance of monitoring 
- Acute and chronic decline associated phenotypes
- The oak decline spectrum move away from simplistic binary categorisations
- Manual assignment of decline status difficult prone to error for due to contribution of many factors. 
- Random forest
- Unsupervised learning for complex patterns between descriptors.

Oak declines are complex syndromes that are facilitated by a variety of both abiotic and biotic factors.
Abiotic factors can include both predisposing and inciting factors such as drought and are often

Biotic factors can include...

As a results of this, there are many visual phenotypic descriptors encompassing the entirety of an oak tree that contribute to what can be defined as a declining oak tree.
A severely declining oak is most prominantly characterised by very poor crown condition and numerous dead branches (ref).

Compounding this complexity is the addition of different types of oak decline (AOD) that are often categorised based on the duration over which a tree has been declining or the biotic agents/factors associated with a declining tree.
For instance a chronically declining oak can subsequently begin to also develop the stem bleeds associated with AOD.
Accurately diagnosing in the field the severity and type of oak decline present is difficult and can requires much experience, expertise and training.

Machine learning is finding increasing utility in biological and geographical applications due to it's ability to learn complex patterns withing and between 
A popular algroith is that of random forest due to it's versatility, relatively trivial training requirements and resiliance against overfitting [@breiman2001] (refs). 
It is an ensemble machine learning method where forests of bootstrap sampled decision trees are grown, yielding both out of bag (OOB) sample prediction accuracies and variable importance.
Unsupervised learning is also possible [@shi2006]

@@

### Phenotypic descriptor preparation

@@

### Machine learning analyses

The code used to generate the decline indexes is available as an R package at [https://github.com/jasenfinch/pdi].
All raw data and code used to generate this manuscript are available at [https://github.com/jasenfinch/Index_measures_for_oak_decline_severity_using_phenotypic_descriptors]
